Projecting genetic associations through gene expression patterns highlights disease etiology and drug mechanisms

### 2023-06-15 output/latex/manuscript.tex
figure 2b, table 2 and 3: provide more details about statistics

\begin{figure}
\hypertarget{fig:lv246}{%
\centering
\includesvg[width=1\textwidth,height=\textheight]{images/lvs_analysis/lv246/lv246.svg}
\caption{\textbf{Tissues and traits associated with a gene module related to lipid metabolism (LV246).}
\textbf{a)} Top cell types/tissues in which LV246's genes are expressed.
Values in the \(y\)-axis come from matrix \(\mathbf{B}\) in the MultiPLIER models (Figure \ref{fig:entire_process}b, see Methods).
In the \(x\)-axis, cell types/tissues are sorted by the maximum sample value.
\textbf{b)} Gene-trait associations (S-MultiXcan; threshold at -log(\(p\))=10) and colocalization probability (fastENLOC) for the top traits in LV246.
The top 40 genes in LV246 are shown, sorted by their LV weight (matrix \(\mathbf{Z}\)), from largest (the top gene \emph{SCD}) to smallest (\emph{FAR2});
\emph{DGAT2} and \emph{ACACA}, in boldface, are two of the six high-confidence genes in the lipids-increasing gene set from the CRISPR screen.
Cardiovascular-related traits are in boldface.
SGBS: Simpson Golabi Behmel Syndrome;
CH2DB: CH2 groups to double bonds ratio;
HDL: high-density lipoprotein;
RCP: locus regional colocalization probability.
}\label{fig:lv246}
}
\end{figure}

@@

\begin{longtable}[]{@{}llll@{}}
\caption{Gene modules (LVs) nominally enriched for the lipids-increasing gene-set from the CRISPR-screen (\emph{P} \textless{} 0.01). LVs significantly aligned with pathways (FDR \textless{} 0.05) from the MultiPLIER models are shown in boldface.
\label{tbl:sup:lipids_crispr:modules_enriched_increase}}\label{tbl:sup:lipids_crispr:modules_enriched_increase}\tabularnewline
\toprule()
Gene module & Lipids gene-set & Leading edge & p-value \\
\midrule()
\endfirsthead
\toprule()
Gene module & Lipids gene-set & Leading edge & p-value \\
\midrule()
\endhead
\textbf{LV246} & increase & \emph{DGAT2}, \emph{ACACA} & 0.0035 \\
LV702 & increase & \emph{ACACA}, \emph{DGAT2} & 0.0046 \\
\textbf{LV607} & increase & \emph{ACACA}, \emph{DGAT2} & 0.0058 \\
LV890 & increase & \emph{ACACA}, \emph{DGAT2} & 0.0067 \\
\textbf{LV74} & increase & \emph{MBTPS1}, \emph{DGAT2} & 0.0078 \\
\textbf{LV865} & increase & \emph{ACACA}, \emph{DGAT2} & 0.0092 \\
LV841 & increase & \emph{ACACA}, \emph{DGAT2} & 0.0096 \\
\bottomrule()
\end{longtable}

\begin{longtable}[]{@{}llll@{}}
\caption{Gene modules (LVs) nominally enriched for the lipids-decreasing gene-set from the CRISPR-screen (\emph{P} \textless{} 0.01). LVs significantly aligned with pathways (FDR \textless{} 0.05) from the MultiPLIER models are shown in boldface.
\label{tbl:sup:lipids_crispr:modules_enriched_decrease}}\label{tbl:sup:lipids_crispr:modules_enriched_decrease}\tabularnewline
\toprule()
Gene module & Lipids gene-set & Leading edge & p-value \\
\midrule()
\endfirsthead
\toprule()
Gene module & Lipids gene-set & Leading edge & p-value \\
\midrule()
\endhead
LV520 & decrease & \emph{FBXW7}, \emph{TCF7L2} & 0.0006 \\
LV801 & decrease & \emph{UBE2J2}, \emph{TCF7L2} & 0.0022 \\
LV512 & decrease & \emph{FBXW7}, \emph{TCF7L2} & 0.0025 \\
\textbf{LV612} & decrease & \emph{PTEN}, \emph{FBXW7} & 0.0036 \\
LV41 & decrease & \emph{PCYT2}, \emph{TCF7L2} & 0.0041 \\
\textbf{LV838} & decrease & \emph{UBE2J2}, \emph{TCF7L2} & 0.0070 \\
LV302 & decrease & \emph{TCF7L2}, \emph{PTEN} & 0.0083 \\
LV959 & decrease & \emph{TCF7L2}, \emph{PTEN} & 0.0092 \\
\bottomrule()
\end{longtable}

>>>>

\begin{figure}
\hypertarget{fig:lv246}{%
\centering
\includesvg[width=1\textwidth,height=\textheight]{images/lvs_analysis/lv246/lv246.svg}
\caption{\textbf{Tissues and traits associated with a gene module related to lipid metabolism (LV246).}
\textbf{a)} Top cell types/tissues in which LV246's genes are expressed.
Values in the \(y\)-axis come from matrix \(\mathbf{B}\) in the MultiPLIER models (Figure \ref{fig:entire_process}b, see Methods).
In the \(x\)-axis, cell types/tissues are sorted by the maximum sample value.
\textbf{b)} Gene-trait associations (unadjusted \(p\)-values from S-MultiXcan\textsuperscript{\protect\hyperlink{ref-1FFzCXo1s}{24}}; threshold at -log(\(p\))=10) and colocalization probability (fastENLOC) for the top traits in LV246.
The top 40 genes in LV246 are shown, sorted by their LV weight (matrix \(\mathbf{Z}\)), from largest (the top gene \emph{SCD}) to smallest (\emph{FAR2});
\emph{DGAT2} and \emph{ACACA}, in boldface, are two of the six high-confidence genes in the lipids-increasing gene set from the CRISPR screen.
Cardiovascular-related traits are in boldface.
SGBS: Simpson Golabi Behmel Syndrome;
CH2DB: CH2 groups to double bonds ratio;
HDL: high-density lipoprotein;
RCP: locus regional colocalization probability.
}\label{fig:lv246}
}
\end{figure}

@@

\begin{longtable}[]{@{}llll@{}}
\caption{Gene modules (LVs) nominally enriched (using FGSEA\textsuperscript{\protect\hyperlink{ref-Z8WXLD67}{51}}) for the lipids-increasing gene-set from the CRISPR-screen (unadjusted \(p\)-values \textless{} 0.01).
LVs significantly aligned with pathways (FDR \textless{} 0.05) from the MultiPLIER models are shown in boldface.
\label{tbl:sup:lipids_crispr:modules_enriched_increase}}\label{tbl:sup:lipids_crispr:modules_enriched_increase}\tabularnewline
\toprule()
Gene module & Lipids gene-set & Leading edge & p-value \\
\midrule()
\endfirsthead
\toprule()
Gene module & Lipids gene-set & Leading edge & p-value \\
\midrule()
\endhead
\textbf{LV246} & increase & \emph{DGAT2}, \emph{ACACA} & 0.0035 \\
LV702 & increase & \emph{ACACA}, \emph{DGAT2} & 0.0046 \\
\textbf{LV607} & increase & \emph{ACACA}, \emph{DGAT2} & 0.0058 \\
LV890 & increase & \emph{ACACA}, \emph{DGAT2} & 0.0067 \\
\textbf{LV74} & increase & \emph{MBTPS1}, \emph{DGAT2} & 0.0078 \\
\textbf{LV865} & increase & \emph{ACACA}, \emph{DGAT2} & 0.0092 \\
LV841 & increase & \emph{ACACA}, \emph{DGAT2} & 0.0096 \\
\bottomrule()
\end{longtable}

\begin{longtable}[]{@{}llll@{}}
\caption{Gene modules (LVs) nominally enriched (using FGSEA\textsuperscript{\protect\hyperlink{ref-Z8WXLD67}{51}}) for the lipids-decreasing gene-set from the CRISPR-screen (unadjusted \(p\)-values \textless{} 0.01).
LVs significantly aligned with pathways (FDR \textless{} 0.05) from the MultiPLIER models are shown in boldface.
\label{tbl:sup:lipids_crispr:modules_enriched_decrease}}\label{tbl:sup:lipids_crispr:modules_enriched_decrease}\tabularnewline
\toprule()
Gene module & Lipids gene-set & Leading edge & p-value \\
\midrule()
\endfirsthead
\toprule()
Gene module & Lipids gene-set & Leading edge & p-value \\
\midrule()
\endhead
LV520 & decrease & \emph{FBXW7}, \emph{TCF7L2} & 0.0006 \\
LV801 & decrease & \emph{UBE2J2}, \emph{TCF7L2} & 0.0022 \\
LV512 & decrease & \emph{FBXW7}, \emph{TCF7L2} & 0.0025 \\
\textbf{LV612} & decrease & \emph{PTEN}, \emph{FBXW7} & 0.0036 \\
LV41 & decrease & \emph{PCYT2}, \emph{TCF7L2} & 0.0041 \\
\textbf{LV838} & decrease & \emph{UBE2J2}, \emph{TCF7L2} & 0.0070 \\
LV302 & decrease & \emph{TCF7L2}, \emph{PTEN} & 0.0083 \\
LV959 & decrease & \emph{TCF7L2}, \emph{PTEN} & 0.0092 \\
\bottomrule()
\end{longtable}
